---
title: "FrequencyFinder4"
output: html_document
date: "2026-08-16"
---

# Setup

## Clean Environment

```{r}
# List of items, mostly large things that are annoying to reinitialise
keep <- c("gtfs") 
# Remove everything that is not in the keep list
rm(list = setdiff(ls(), keep))
# Get rid of everything if needed
#rm(list = ls())
```

## Load Libraries

```{r}
library(data.table)
library(sf)
library(jsonlite)
library(gtfstools)
library(vicgtfstools) # My own repo https://github.com/adambain014/vicgtfstools 
library(tictoc) # For my own benchmarking enjoyment
library(ggplot2) # For diagnostics
```

## Initialise the GTFS Datasets

```{r}
# Create a temps folder
if(!dir.exists("Temps")){
  dir.create("Temps")
  }

# Download the latest vic gtfs feed
if(!dir.exists("Temps/gtfs")){
  download_latest_vic_gtfs("Temps/gtfs")
}

# Open the GTFS
if(!exists("gtfs")){
  gtfs <- open_and_join_all_vic_gtfs("Temps/gtfs") |> tag_myki_bus()
}
```

## Select the modes I will be working on - it will be expanded
Challenge: inconsistent naming system

```{r}
select_modes <- c(1,2,3,4,10,11) # V/Line. Metro train, tram and bus
```

# Data processing 

## Create a representative stop times of a 1 week period - this will be unforuntatley large 
```{r}
# Open the calendar file
calendar <- gtfs$calendar[mode_number %in% select_modes]

# Extract entries that are within the first full week
first_date <- min(calendar[weekdays(start_date) == "Monday" & start_date > min(gtfs$calendar$start_date) + 7]$start_date)
last_date <- first_date + 6
calendar_first_full_week <- calendar[start_date <= last_date & end_date >= first_date]


# Melt into long format
calendar_first_full_week_long <- melt(
  calendar_first_full_week,
  id.vars = c(1, 9:12),
  measure.vars = 2:8,
  variable.name = "day",
  value.name = "bool"
)

# Add day number (0 for monday etc)
calendar_first_full_week_long[
  , day_number :=
      fifelse(day == "monday", 0,
      fifelse(day == "tuesday", 1,
      fifelse(day == "wednesday", 2,
      fifelse(day == "thursday", 3,
      fifelse(day == "friday", 4,
      fifelse(day == "saturday", 5,
      fifelse(day == "sunday", 6, NA_integer_)))))))]

# Add date entry
calendar_first_full_week_long[, date := first_date + day_number]

# Deal with calendar dates
calendar_dates <- gtfs$calendar_dates[mode_number %in% select_modes]

calendar_first_full_week_long[
  , exception_type := calendar_dates[.SD, on = .(date, service_id, mode_number), exception_type]
]

calendar_first_full_week_long[
  , bool := fifelse(date < start_date | date > end_date, 0L, bool)
]

# exception_type meanings:
# 1 = added service, does not exist in VIC GTFS
# 2 = removed service
# NA = no exception

calendar_first_full_week_long[!is.na(exception_type)
  , bool := fifelse(exception_type == 2, 0,
           fifelse(exception_type == 1, 1, bool))
]

# Simplify the calendar, remove service_ids from days that are not needed.
calendar_refined <- calendar_first_full_week_long[bool == 1, .(service_id, day_number, mode_number)]

# Attach this information to trips
trips <- gtfs$trips[mode_number %in% select_modes]
trips <- trips[calendar_refined, on = .(mode_number, service_id), nomatch = 0L, allow.cartesian = TRUE]

# Clean up calendar intermediaries
rm(calendar_first_full_week, calendar_first_full_week_long, calendar_refined)

# Add route_short_name from routes
routes <- gtfs$routes[mode_number %in% select_modes]

# Pick off replacement buses and ANZAC day busrs
routes_refined <- routes[
  !route_short_name %chin% c("Replacement Bus", "ANZAC_1", "ANZAC_2", "ANZAC_3")]

# Fix Bacchus Marsh - will have to fix in vicgtfstools when I get round to a review
routes_refined[route_short_name == "", route_short_name := tstrsplit(route_id, "-")[2]]
routes_refined[mode_number == 4 , route_code := tstrsplit(route_id, "-", fill = NA)[2]]
routes_refined[mode_number == 4 & grepl("^BM",route_code), region := "Bacchus Marsh"]
routes_refined[mode_number == 4 & grepl("^WN",route_code), region := "Wallan"]

# Add suffix for regional routes
routes_refined[mode_number == 4 &
                 region != "Melbourne", 
               route_short_name := paste0(route_short_name, " (", region, ")")]

# Fix route colour for the overland
routes_refined[mode_number == 10, route_color := "#140075"]

# For sky bus use official name
sb_map <- data.table(
  mode_number = c(11,11,11,11,11),
  route_long_name = c("Box Hill - Melbourne Airport",
                      "Frankston - Melbourne Airport",
                      "Melbourne City - Avalon Airport",
                      "City - Melbourne Airport",
                      "Sunshine Railway Station - Melbourne Airport"),
  route_short_name = c("Eastern Express",
                       "Peninsula Express",
                       "Avalon City Express",
                       "Melbourne City Express",
                       "Sunshine Express")
)

routes_refined[sb_map, on = .(mode_number, route_long_name), route_short_name := i.route_short_name]


routes_refined <- routes_refined[,.(route_id, mode_number, route_short_name, route_color)]

# Add to trips
trips <- trips[route_id %chin% routes_refined$route_id]
trips <- routes_refined[trips, on = .(mode_number, route_id)]

# Condense trips
trips_refined <- trips[, .(trip_id, route_short_name, day_number, mode_number, direction_id)]

# Clean up intermediaries
rm(routes_refined, sb_map)

# Open stop_times
stop_times <- gtfs$stop_times[mode_number %in% select_modes]

# Keep only needed columns
cols_to_keep <- c(
  "trip_id",
  "departure_time",
  "stop_id",
  "stop_sequence",
  "shape_dist_traveled",
  "mode_number"
)
stop_times <- stop_times[, ..cols_to_keep]
rm(cols_to_keep)

stop_times <- stop_times[trip_id %chin% trips$trip_id]

# Join onto stop_times
stop_times <- trips[,.(trip_id, route_short_name, day_number, mode_number, direction_id, route_color)
                    ][stop_times, on = .(trip_id, mode_number), allow.cartesian = TRUE]

rm(trips_refined)
# remove dupes
stop_times <- unique(stop_times, by = 
                           c("stop_id",
                             "departure_time",
                             "direction_id",
                             "route_short_name",
                             "shape_dist_traveled",
                             "day_number",
                             "mode_number"))

# Turn departure times into seconds
stop_times[, c("h","m","s") := tstrsplit(departure_time, ":", type.convert = TRUE)]
stop_times[, departure_time_m := round(h*60 + m + s/60)]
stop_times <- stop_times[, !c("h","m","s")]

# Mark for past midnight
stop_times <- stop_times[,past_midnight := departure_time_m >= 24 * 60]

stops <- gtfs$stops[mode_number %in% select_modes]

# Add station code to deal with multi-platform train stations
parent_lookup <- unique(
  stops[mode_number == 2 & !is.na(parent_station) & parent_station != "",
        .(stop_id, mode_number, parent_station)]
)

stop_times <- parent_lookup[stop_times, on = c("stop_id", "mode_number")]
stop_times[!is.na(parent_station), stop_id := parent_station]
stop_times <- stop_times[, -"parent_station"]

rm(parent_lookup)

# Fix 790 inconsistent direction_id by flipping it
stop_times[route_short_name == 790, direction_id := fifelse(direction_id == 0, 1, 0)]
# and 426
stop_times[route_short_name == 426, direction_id := fifelse(direction_id == 0, 1, 0)]
# and 283
stop_times[route_short_name == 281, direction_id := fifelse(direction_id == 0, 1, 0)]

stop_times[mode_number==10]
```


## Add combined corridors
```{r}
corridor_defs <- list(
  # Train
  "Belgrave-Lilydale"     = c("Belgrave", "Lilydale"),
  "Cranbourne-Pakenham"   = c("Cranbourne", "Pakenham"),
  "Hurstbridge-Mernda"    = c("Hurstbridge", "Mernda"),
  "Werribee-Williamstown" = c("Werribee", "Williamstown"),

  # Tram
  "1-6"                 = c("1", "6"),
  "1-3-5-6-16-64-67-72" = c("1", "3", "5", "6", "16", "64", "67", "72"),
  "5-64"                = c("5", "64"),
  "12-109"              = c("12", "109"),

  # Bus
  "200-207" = c("200", "207"),
  "250-251" = c("250", "251"),
  "281-293" = c("281", "293"),
  "302-304" = c("302", "304"),
  "303-302-304-305-309-318-350-905-906-907-908" =
    c("302", "303", "304", "305", "309", "318", "350", "905", "906", "907", "908"),
  "386-387" = c("386", "387"),
  "411-412" = c("411", "412"),
  "426-456" = c("426", "456"),
  "478-479" = c("478", "479"),
  "513-514" = c("513", "514"),
  "782-783" = c("782", "783"),
  "781-784-785" = c("781", "784", "785"),
  "789-790-791" = c("789", "790", "791"),
  "802-804-862" = c("802", "804", "862"),
  "811-812" = c("811", "812"),
  "814-848" = c("814", "848"),
  "843-845" = c("843", "845"),
  "600-922-923" = c("600", "922", "923")
)


combine_corridors <- function(st, corridor_defs){
  # corridor_defs: named list, e.g. list("1/6" = c("1","6"), "3/64/67" = c("3","64","67"))
  # name = new combined route_short_name, value = character vector of routes to merge
  
  setkey(st, route_short_name)  # enables fast binary-search subsetting below instead of full scans
  
  # Remove existing combined corridors
  st <- st[!route_short_name %chin% names(corridor_defs)]
  
  combined_list <- lapply(names(corridor_defs), function(new_name){
    routes <- corridor_defs[[new_name]]
    
    # One binary-search subset per route (fast, keyed lookup rather than vector scan)
    subsets <- lapply(routes, function(r) st[.(r)])
    
    # Stops common to every route in the group - the shared corridor segment
    shared_stops <- Reduce(intersect, lapply(subsets, function(x) x$stop_id))
    
    combined <- rbindlist(lapply(subsets, function(x) x[stop_id %chin% shared_stops]))
    combined[, route_short_name := new_name]
    combined
  })
  
  # Single rbind at the end, not one per corridor - avoids repeatedly copying the growing table
  rbind(st, rbindlist(combined_list))
}


stop_times <- combine_corridors(stop_times, corridor_defs)
```

# Subdivide corridors - for V/Line
```{r}
# Subdivide V/Line - Geelong into 3 - Warrnambool, Wyndham Vale, Geelong,
# Long distance is easy as it an be inferred from trip_id
# Short runs will have to be duplicated from the leftovers,
# Then short run only trips must be deleted from the whole line
# I might consider doing this with the Lilydale line because the Blackburn stoppers are leading to some dishonest results
ld_map <- data.table(
  pattern = c("WBL",
              "ART",
              "MBY",
              "SWL",
              "ECH",
              "SNH",
              "ABY",
              "BDE"),
  route_short_name = c("Warrnambool",
                       "Ararat",
                       "Maryborough",
                       "Swan Hill",
                       "Echuca",
                       "Shepparton",
                       "Albury",
                       "Bairnsdale")
)

stop_times[mode_number == 1, prefix := tstrsplit(trip_id, "-", fixed = TRUE)[2]]
stop_times[ld_map, on = .(prefix = pattern),
           route_short_name := i.route_short_name]

rm(ld_map)

# Deal with the short suburban runs
suburban_short_run <- function(st, stop_list, old_group_name, new_name){
  stop_ids <- stops[stop_name %chin% stop_list]$stop_id |> unique()
  short_run_st <- st[route_short_name == old_group_name & stop_id %chin% stop_ids]
  short_run_st[, route_short_name := new_name]
  rbind(st, short_run_st)
}

# List of stops for each
wyndham_vale_stops <- c("Wyndham Vale Station",
                        "Tarneit West Station",
                        "Tarneit Station",
                        "Truganina Station",
                        "Deer Park Station",
                        "Sunshine Station",
                        "Footscray Station",
                        "Southern Cross Station")

melton_stops       <- c("Melton Station",
                        "Cobblebank Station",
                        "Thornhill Park Station",
                        "Rockbank Station",
                        "Mount Atkinson Station",
                        "Caroline Springs Station",
                        "Deer Park Station",
                        "Ardeer Station",
                        "Sunshine Station",
                        "Footscray Station",
                        "Southern Cross Station")

stop_times <- suburban_short_run(stop_times, wyndham_vale_stops, "Geelong", "Wyndham Vale")
stop_times <- suburban_short_run(stop_times, melton_stops,      "Ballarat", "Melton")

# Clean up the overlap
geelong_stop_id <- stops[stop_name == "Geelong Station"]$stop_id
geelong_trips <- stop_times[stop_id == geelong_stop_id]$trip_id
stop_times <- stop_times[!route_short_name == "Geelong" | trip_id %chin% geelong_trips]

b_bm_stop_ids <- stops[stop_name == "Bacchus Marsh Station" | stop_name == "Ballarat Station"]$stop_id
ballarat_trips <- stop_times[stop_id %chin% b_bm_stop_ids]$trip_id |> unique()
stop_times <- stop_times[!route_short_name == "Ballarat" | trip_id %chin%  ballarat_trips]

rm(suburban_short_run, 
   wyndham_vale_stops, 
   geelong_stop_id, 
   melton_stops, 
   geelong_trips, 
   b_bm_stop_ids, 
   ballarat_trips)
```

# Function for inspection (in progress)
```{r, eval=FALSE}
plot_stop_times <- function(filtered_st, stop_order) {
  
  plot_dt <- filtered_st[!is.na(departure_time_m)]
  
  ggplot(plot_dt, aes(x = departure_time_m, y = label, group = trip_id)) +
    geom_line(alpha = 0.4, colour = "steelblue") +
    geom_point(size = 0.8, alpha = 0.6, colour = "steelblue") +
    scale_y_discrete(limits = rev(levels(plot_dt$label))) +  # top stop at top
    scale_x_continuous(
      breaks = seq(0, 1440, by = 60),
      labels = function(m) sprintf("%02d:%02d", m %/% 60, m %% 60)
    ) +
    labs(x = "Time", y = NULL, title = "Trip diagram") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank())
}

inspect_stop_times <- function(st, rsn, dn, did, plot = FALSE){
  
  filtered_st <- gtfs$stops[,.(stop_id, stop_name)][
      st[route_short_name == rsn & day_number == dn & direction_id == did],
      on = "stop_id"
  ]
  
  # find the most common stop_id sequence among trips
    trip_patterns <- filtered_st[order(stop_sequence),
      .(pattern = paste(stop_id, collapse = "|"),
        n_stops = .N), by = trip_id]
  
  pattern_summary <- trip_patterns[, .(n_stops = n_stops[1], n_trips = .N), by = pattern]
  longest_pattern <- pattern_summary[order(-n_stops, -n_trips)]$pattern[1]
  template_trip <- trip_patterns[pattern == longest_pattern, trip_id][1]
  
  base_order <- filtered_st[
      trip_id == template_trip
  ][order(stop_sequence), .(stop_id, base_rank = .I)]
  
  # median departure time per stop, used to place non-base stops
  stop_time <- filtered_st[
      , .(med_dep = median(departure_time_m, na.rm = TRUE)), by = .(stop_id, stop_name)
  ]
  
  stop_order <- base_order[stop_time, on = "stop_id"]
  
  # insert stops missing from the template between their nearest base neighbours by time
  base_times <- stop_order[!is.na(base_rank)][order(base_rank), .(base_rank, med_dep)]
  stop_order[is.na(base_rank), base_rank := sapply(med_dep, function(t) {
      idx <- which.min(abs(base_times$med_dep - t))
      base_times$base_rank[idx] + ifelse(t < base_times$med_dep[idx], -0.5, 0.5)
  })]
  
  stop_order <- stop_order[order(base_rank, med_dep)]
  stop_order[, label := paste0(sprintf("%03d", .I), " ", stop_name)]
  
  filtered_st[stop_order, on = "stop_id", label := i.label]
  filtered_st[, label := factor(label, levels = stop_order$label)]
  
    if (plot) {
    print(plot_stop_times(filtered_st, stop_order))
  }
  
  trip_start <- filtered_st[
      , .(trip_first_dep = min(departure_time_m, na.rm = TRUE)), by = trip_id
  ][order(trip_first_dep)]
  
  out <- dcast(filtered_st, label ~ trip_id, value.var = "departure_time",
               fun.aggregate = function(x) x[1])
  
  setcolorder(out, c("label", as.character(trip_start$trip_id)))
  out
}
```

# Summary Stats
Average Frequency for Morning, Daytime, and Evening across Monday, Saturday and Sunday
```{r}

# Add minutes until next service in same direction at same stop
waits <- copy(stop_times)

# Sort to ensure correct order
setorder(waits, mode_number, day_number, route_short_name,  direction_id, stop_id, departure_time_m)

# Find next service
waits[,
  next_time_m := shift(departure_time_m, 1L, type = "lead"),
  by = .(stop_id, direction_id, day_number, mode_number, route_short_name)
]

# Get rid of last services
waits <- waits[!is.na(next_time_m)]

waits[, wait_m := next_time_m - departure_time_m]

test <- stops[, .(stop_id, stop_name)][
  waits[route_short_name == "Sandringham" & day_number == 0 & direction_id == 1 & 
          departure_time_m > 7 * 60 & departure_time_m <= 8 * 60],
  on = "stop_id"
][order(trip_id, stop_sequence)]


```

```{r}
# --------------------------------------------------------------------------
# Build per-route frequency summary across a set of day/time windows
# --------------------------------------------------------------------------

# Skeleton table of every route present in `waits` - final results get
# left-joined onto this so routes with no service in a window still appear
# (with NA rather than being dropped from the output entirely)
routes_list <- data.table(route_short_name = unique(waits$route_short_name))
## Add the overland
routes_list <- rbind(routes_list, data.table(route_short_name = "The Overland"))

# Computes average headway (mean wait between consecutive services) for one
# route, over one day_number, within one clock-time window.
#
# Method: mean-of-means, not a single pooled mean.
#   1. Average wait_m separately for every (route, stop_id, direction_id)
#      combo in the window - this is one "headway estimate" per stop+direction.
#   2. Average those stop-level means up to a single route-level figure.
# This gives every stop along the route equal weight regardless of how many
# individual departures pass through it, so busy interchange stops don't
# dominate the route's headway figure. Deliberate choice - note this in any
# write-up, since it differs from simply averaging every wait_m row directly.
freq_at_window_dt <- function(waits, target_day, start_min, end_min, cap_min = 120) {
  
  # Filter to the target day and clock-time window.
  # NB: departure_time_m is in MINUTES since midnight, not seconds - start_min/
  # end_min must be too (see windows table below, uses `* 60`).
  window_waits <- waits[
    day_number == target_day &
    departure_time_m >= start_min &
    departure_time_m <  end_min
  ]
  
  # No services at all in this window/day -> return an empty (typed) table
  # rather than NULL, so rbindlist() downstream doesn't choke on missing cols
  if (nrow(window_waits) == 0) {
    return(data.table(route_short_name = character(), mean_wait = numeric()))
  }
  
  # Cap absurdly long waits (e.g. last service of the day, or gaps from
  # low-frequency weekend running) so they don't blow out the route average.
  # Anything >= cap_min (default 2 hrs) is treated as missing rather than
  # included as a genuine "wait".
  window_waits[wait_m >= cap_min, wait_m := NA]
  
  out <- window_waits[
    # Step 1: mean wait per stop+direction (one headway estimate per stop)
    , .(mean_wait = mean(wait_m, na.rm = TRUE)),
    by = .(route_short_name, stop_id, direction_id)
  ][
    # Step 2: average those stop-level means up to one figure per route
    , .(mean_wait = mean(mean_wait, na.rm = TRUE)),
    by = route_short_name
  ]

  # Round to nearest half-minute for a cleaner, less falsely-precise output
  out[, mean_wait := round(mean_wait * 2) / 2]
  out[]
}

# Reference windows: three time-of-day slices (AM peak / midday / evening)
# across three representative days (Monday = weekday, Saturday, Sunday).
# day_number: 0 = Monday ... 6 = Sunday (see day_number derivation earlier
# in the script). start_min/end_min are minutes-since-midnight (7*60 = 7am).
windows <- data.table(
  window_name = c("mon_7_8", "mon_12_13", "mon_19_20", 
                  "sat_7_8", "sat_12_13", "sat_19_20",
                  "sun_7_8", "sun_12_13", "sun_19_20"),
  day_number  = c(0,        0,        0,        5,        5,        5,        6,        6,        6),
  start_min   = c(7*60,     12*60,    19*60,    7*60,     12*60,    19*60,    7*60,     12*60,    19*60),
  end_min     = c(8*60,     13*60,    20*60,    8*60,     13*60,    20*60,    8*60,     13*60,    20*60)
)

# Run freq_at_window_dt once per window, tag each result with its window_name,
# and stack into one long-format table (route_short_name x window_name x mean_wait)
freq_results <- rbindlist(lapply(seq_len(nrow(windows)), function(i) {
  w   <- windows[i]
  res <- freq_at_window_dt(waits, w$day_number, w$start_min, w$end_min)
  if (nrow(res) > 0) res[, window_name := w$window_name]
  res
}))

# Pivot long -> wide: one row per route, one column per window
freq_wide <- dcast(freq_results, route_short_name ~ window_name, value.var = "mean_wait")

# Form for JSON formation
freq_wide[, frequency_summary := Map(
  function(m7, m12, m19, s7, s12, s19, su7, su12, su19) {
    list(
      list(morning = m7,  daytime = m12, evening = m19, day = "Monday"),
      list(morning = s7,  daytime = s12, evening = s19, day = "Saturday"),
      list(morning = su7, daytime = su12, evening = su19, day = "Sunday")
    )
  },
  mon_7_8, mon_12_13, mon_19_20,
  sat_7_8, sat_12_13, sat_19_20,
  sun_7_8, sun_12_13, sun_19_20
)]


# Clean existing routes
to_remove <- windows$window_name[windows$window_name %chin% names(routes_list)]
routes_list[,(to_remove) := NULL]
rm(to_remove)

# Left-join onto the full route list so every route appears in the final
# output even if it had no service (and therefore no result row) in every window
routes_list <- merge(routes_list, freq_wide, by = "route_short_name", all.x = TRUE)
```

# Diagnostic chart

```{r, eval=FALSE}
freq_histo <- function(rl, # routes_list
                       period,
                       sub){
  ggplot(
    routes_list[!routes_list$route_short_name %chin% names(corridor_defs) & !is.na(get(period))],
    aes(x = get(period))
  ) +
    geom_histogram(
      binwidth = 1,
      fill = "grey40",
      color = "white",
      alpha = 0.9
    ) +
    scale_y_continuous(limits = c(0, 110)) +
    labs(
      x = "Average Wait (minutes)",
      y = "Number of Routes",
      title = "Distribution of Average Wait Times",
      subtitle = sub,
      caption = "Chart by Adam Bain, Data from DTP 2026"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    )
}

freq_histo(routes_list, "mon_7_8", "Monday Peak (7am-8am)")
freq_histo(routes_list, "mon_12_13", "Monday Daytime (12pm-1pm)")
freq_histo(routes_list, "sat_12_13", "Saturday Daytime (12pm-1pm)")
freq_histo(routes_list, "sun_12_13", "Sunday Daytime (12pm-1pm)")

```

## Peak Medians
```{r}
median_in_window_dt <- function(waits, target_day, start_min, end_min, cap_min = 120) {
  
  # Filter to the target day and clock-time window.
  # NB: departure_time_m is in MINUTES since midnight, not seconds - start_min/
  # end_min must be too (see windows table below, uses `* 60`).
  window_waits <- waits[
    day_number == target_day &
    departure_time_m >= start_min &
    departure_time_m <  end_min
  ]
  
  # No services at all in this window/day -> return an empty (typed) table
  # rather than NULL, so rbindlist() downstream doesn't choke on missing cols
  if (nrow(window_waits) == 0) {
    return(data.table(route_short_name = character(), mean_wait = numeric()))
  }
  
  # Cap absurdly long waits (e.g. last service of the day, or gaps from
  # low-frequency weekend running) so they don't blow out the route average.
  # Anything >= cap_min (default 2 hrs) is treated as missing rather than
  # included as a genuine "wait".
  window_waits[wait_m >= cap_min, wait_m := NA]
  
  out <- window_waits[
    # Median
    , .(median_wait = median(wait_m, na.rm = TRUE)),
    by = .(route_short_name)
  ]
  out[]
}

# Reference windows: AM and PM peak - 7am to 9am and 4pm to 7pm
windows <- data.table(
  window_name = c("am_peak", "pm_peak"),
  day_number  = c(0,        0), 
  start_min   = c(7*60,     16*60),
  end_min     = c(9*60,     19*60)  
)

# Run median_in_window_dt once per window, tag each result with its window_name,
# and stack into one long-format table (route_short_name x window_name x mean_wait)
freq_results <- rbindlist(lapply(seq_len(nrow(windows)), function(i) {
  w   <- windows[i]
  res <- median_in_window_dt(waits, w$day_number, w$start_min, w$end_min)
  if (nrow(res) > 0) res[, window_name := w$window_name]
  res
}))

# Pivot long -> wide: one row per route, one column per window
freq_wide <- dcast(freq_results, route_short_name ~ window_name, value.var = "median_wait")


# Form for JSON formation
freq_wide[, peak_medians := Map(
  function(am, pm) {
    list(
      list(
        "Median AM Peak Wait" = am,
        "Median PM Peak Wait" = pm
      )
    )
  },
  am_peak,
  pm_peak
)]


# Clean existing routes
to_remove <- windows$window_name[windows$window_name %chin% names(routes_list)]
routes_list[,(to_remove) := NULL]
rm(to_remove)

# Left-join onto the full route list so every route appears in the final
# output even if it had no service (and therefore no result row) in every window
routes_list <- merge(routes_list, freq_wide, by = "route_short_name", all.x = TRUE)


```

## Span chart
```{r}
  # Guard against infinites caused by missing days
safe_max <- function(x) if (length(x) == 0 || all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
safe_min <- function(x) if (length(x) == 0 || all(is.na(x))) NA_real_ else min(x, na.rm = TRUE)
  
## Span stats (vectorised, whole-table version)
# Same logic as FF3's one_day_span/span_across_week, but computed once
# across every route+day simultaneously via grouped data.table operations,
# consistent with FF4's whole-network approach rather than per-route calls.

# --- Step 1: valid stops per route+day ------------------------------------
# Drop stops served by <20% of that route+day's peak stop-visit count
# (depot-only/rare terminus variants), same threshold as FF3, but computed
# for every route+day combo at once.
stop_counts <- stop_times[, .N, by = .(route_short_name, day_number, stop_id)]
stop_counts[, pc_max := N / max(N), by = .(route_short_name, day_number)]
valid_stops <- stop_counts[pc_max > 0.2, .(route_short_name, day_number, stop_id)]

# Semi-join stop_times down to only valid stops, per route+day
stop_times_valid <- stop_times[valid_stops, on = .(route_short_name, day_number, stop_id), nomatch = 0L]

# --- Step 2: span_start/span_end per route+day -----------------------------
span_tbl <- stop_times_valid[
  , .(span_start = safe_min(departure_time_m),
      span_end   = safe_max(departure_time_m)),
  by = .(route_short_name, day_number)
]

# --- Step 3: full coverage window per route+day -----------------------------
# First/last service per stop, then the tightest common window across stops
cov_by_stop <- stop_times_valid[
  , .(first_service = safe_min(departure_time_m),
      last_service  = safe_max(departure_time_m)),
  by = .(route_short_name, day_number, stop_id)
]

cov_tbl <- cov_by_stop[
  , .(full_coverage_start = safe_max(first_service),
      full_coverage_end   = safe_min(last_service)),
  by = .(route_short_name, day_number)
]

# --- Step 4: assemble, including route+day combos with zero service --------
# CJ gives the full grid so routes with no service on a given day still get
# a row (with NA), matching FF3's explicit per-day handling
all_combos <- CJ(route_short_name = unique(stop_times$route_short_name), day_number = 0:6)

span_summary <- span_tbl[cov_tbl, on = .(route_short_name, day_number)]
span_summary <- all_combos[span_summary, on = .(route_short_name, day_number)]  # ensure full grid even if either side was missing combos

# --- Step 5: night network detection, vectorised per route -----------------
# For each route, rotate day_number 0-6 so Sunday's late running can connect
# to Monday's early start (c(.N,1:(.N-1)) / c(2:.N,1) tricks from FF3, now
# applied per-group across the whole table via `by`)
setorder(span_summary, route_short_name, day_number)

span_summary[, span_end_n := span_end - 24 * 60]

span_summary[
  , prior_span_end_n := span_end_n[c(.N, seq_len(.N - 1))],
  by = route_short_name
]

span_summary[, span_diff            := prior_span_end_n - span_start]
span_summary[, begins_night_network := span_diff >= 30]

span_summary[
  , ends_night_network := begins_night_network[c(seq_len(.N)[-1], 1)],
  by = route_short_name
]

span_summary[begins_night_network == TRUE, span_start := 3 * 60]
span_summary[begins_night_network == TRUE, full_coverage_start := 3 * 60]

span_summary[ends_night_network == TRUE, span_end := 27 * 60]
span_summary[ends_night_network == TRUE, full_coverage_end := 27 * 60]

# --- Step 6: tidy up ---------------------------------------------------------
day_names <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
span_summary[, day := day_names[day_number + 1]]

span_summary <- span_summary[, .(route_short_name, day, day_number, span_start, span_end,
                                  full_coverage_start, full_coverage_end)]
setorder(span_summary, route_short_name, day_number)


## Contort span_summary into pre-JSON nested form, join to routes_list

day_names_lower <- c("monday","tuesday","wednesday","thursday","friday","saturday","sunday")

# Build one row per route_short_name, with a list-of-7-lists column
# (one list per day, in Monday->Sunday order) matching the frequency_summary/
# peak_medians nesting pattern already used above
span_nested <- span_summary[
  , .(span_stats = list(
      lapply(0:6, function(d){
        row <- .SD[day_number == d]
        list(
          day                  = day_names_lower[d + 1],
          span_start           = row$span_start,
          span_end             = row$span_end,
          full_coverage_start  = row$full_coverage_start,
          full_coverage_end    = row$full_coverage_end
        )
      })
    )),
  by = route_short_name
]

# Left-join onto routes_list, same convention as frequency_summary/peak_medians
routes_list <- merge(routes_list, span_nested, by = "route_short_name", all.x = TRUE)

# Clean up intermediaries
rm(stop_counts, valid_stops, stop_times_valid, span_tbl, cov_by_stop, cov_tbl, all_combos)


```
```{r}

group_sizes <- span_summary[, .N, by = route_short_name]
setorder(group_sizes, route_short_name)
group_sizes[416]        # this is your crashing group specifically
group_sizes[N != 7]     # this shows you every problem route, not just #416
```

## Add route data - weekly km, hours, trips and avg speed
```{r}
weekly_route_data <- function(stop_times) {
  st <- copy(stop_times)
  setorder(st, trip_id, day_number, stop_sequence)

  st_by_trip_day <- st[, .(
    trip_length_m = shape_dist_traveled[.N] - shape_dist_traveled[1],
    trip_time_min = departure_time_m[.N] - departure_time_m[1]
  ), by = .(trip_id, day_number, route_short_name)]

  st_by_rsn_weekly <- st_by_trip_day[, .(
    trip_count = .N,
    length_km  = round(sum(trip_length_m) / 1000),
    time_hr    = round(sum(trip_time_min) / 60)
  ), by = route_short_name]

  st_by_rsn_weekly[, avg_speed_kmhr := fifelse(time_hr > 0, round(length_km / time_hr, 2), NA_real_)]
  st_by_rsn_weekly[]
}

weekly_dt <- weekly_route_data(stop_times)

weekly_dt[, weekly_stats := Map(
  function(tc, km, hr, sp) {
    list(
      list(
        "Weekly Trip Count"      = tc,
        "Weekly Distance (km)"   = km,
        "Weekly Time (hr)"       = hr,
        "Average Speed (km/hr)"  = sp
      )
    )
  },
  trip_count,
  length_km,
  time_hr,
  avg_speed_kmhr
)]

# Clean existing routes
to_remove <- setdiff(names(weekly_dt), "route_short_name")
routes_list[,(to_remove) := NULL]
rm(to_remove)


routes_list <- merge(routes_list, weekly_dt[,.(route_short_name, weekly_stats)], 
                     by = "route_short_name", all.x = TRUE)
```


## Add metadata for each route and export summary

```{r}
combined_corridors <- names(corridor_defs)

summary <- routes_list[,.(route_short_name,
                          frequency_summary,
                          peak_medians,
                          span_stats,
                          weekly_stats)]

# Add mode number
summary <- unique(waits[,.(route_short_name, mode_number)])[summary, on = "route_short_name"]
summary[route_short_name == "The Overland", mode_number := 10]
# Create fake reference route_short_name for the combined corridors
summary[, route_number := route_short_name]
summary[route_short_name %chin% combined_corridors, 
        route_short_name := sapply(route_number, function(x) corridor_defs[[x]][1])]

# Add colour and mode
summary <- unique(stop_times[,.(route_short_name, route_color, mode_number)
                   ])[summary, on = .(route_short_name, mode_number)]

get_destinations <- function(dir_id) {
  singles <- trips[
    direction_id == dir_id, .N, by = .(route_short_name, trip_headsign)
  ][order(route_short_name, -N)][, .(most_common_hs = trip_headsign[1]), by = route_short_name]

  combined <- stops[,.(stop_id, stop_name)][
    stop_times[route_short_name %chin% combined_corridors], on = "stop_id"
  ][order(trip_id, stop_sequence)
  ][direction_id==dir_id, .(final=last(stop_name)), by=.(route_short_name, trip_id)
  ][, .N, by=.(route_short_name, final)][order(route_short_name,-N)
  ][, .(most_common_hs = final[1]), by = route_short_name]

  out <- rbind(singles, combined)

  # Fallback for my artificial vline routes
  missing_routes <- setdiff(unique(stop_times$route_short_name), out$route_short_name)
  if (length(missing_routes) > 0) {
    fallback <- stops[,.(stop_id, stop_name)][
      stop_times[route_short_name %chin% missing_routes], on = "stop_id"
    ][order(trip_id, stop_sequence)
    ][direction_id == dir_id, .(final = last(stop_name)), by = .(route_short_name, trip_id)
    ][, .N, by = .(route_short_name, final)][order(route_short_name, -N)
    ][, .(most_common_hs = final[1]), by = route_short_name]
    out <- rbind(out, fallback)
  }

  out[, most_common_hs := gsub("Railway Station", "Station", most_common_hs)]
  setnames(out, c("route_short_name", "most_common_hs"),
           c("route_number", paste0("route_destination_", dir_id)))
  out[]
}


# Add direction destinations
rd1 <- get_destinations(1)
rd0 <- get_destinations(0)
summary <- rd1[rd0, on = "route_number"][summary, on = "route_number"]

# Fix order + rename to match FF3's output field names
summary[, "route_short_name" := NULL]
setnames(summary, c("mode_number", "route_color"), c("mode_type", "route_colour"))
summary[, mode_type := as.character(mode_type)]
# Add marker for combined corridor
summary[, is_combined_corridor := fifelse(route_number %chin% combined_corridors, TRUE, FALSE)]
# Display name
summary[, route_display := gsub("-", "/", route_number)] 

# Lookup table for specific display name changes
route_display_changes <- c("1-3-5-6-16-64-67-72" = "Swanston/St Kilda Corridor",
                              "303-302-304-305-309-318-350-905-906-907-908" = "Eastern Freeway Corridor")

summary[route_number %chin% names(route_display_changes),
        route_display := route_display_changes[route_number]]

rm(route_display_changes)
# Short bullet name
summary[, route_bullet_code := route_display]

# Lookup table
route_bullet_changes <- c(
  # Large corridors
  "Swanston/St Kilda Corridor" = "Sw/StK",
  "Eastern Freeway Corridor"   = "M3",
  # Bus pairs / groups
  "200/207" = "200/7",
  "250/251" = "250/1",
  "281/293" = "281/93",
  "302/304" = "302/4",
  "386/387" = "386/7",
  "411/412" = "411/2",
  "426/456" = "426/56",
  "478/479" = "478/9",
  "513/514" = "513/4",
  "600/922/923" = "600/922/3",
  "781/784/785" = "781/4/5",
  "782/783" = "782/3",
  "789/790/791" = "789/90/1",
  "802/804/862" = "802/4/62",
  "811/812" = "811/2",
  "814/848" = "814/48",
  "843/845" = "843/5",
  # Rail lines (standard codes)
  "Alamein"              = "ALM",
  "Belgrave"             = "BEL",
  "Belgrave/Lilydale"    = "BEL/LIL",
  "Craigieburn"          = "CGB",
  "Cranbourne"           = "CBE",
  "Cranbourne/Pakenham"  = "CBE/PKM",
  "Frankston"            = "FKN",
  "Glen Waverley"        = "GWY",
  "Hurstbridge"          = "HBE",
  "Hurstbridge/Mernda"   = "HBE/MDD",
  "Lilydale"             = "LIL",
  "Pakenham"             = "PKM",
  "Sandringham"          = "SHM",
  "Stony Point"          = "STY",
  "Sunbury"              = "SUN",
  "Upfield"              = "UPF",
  "Werribee"             = "WER",
  "Werribee/Williamstown"= "WER/WIL",
  "Williamstown"         = "WIL",
  # V/Line
  "Geelong"      = "GLG",
  "Ballarat"     = "BAT",
  "Bendigo"      = "BGO", 
  "Seymour"      = "SER",
  "Traralgon"    = "TRN",
  "Warrnambool"  = "WBL",
  "Ararat"       = "ART",
  "Maryborough"  = "MYB",
  "Swan Hill"    = "SWL",
  "Echuca"       = "ECH",
  "Shepparton"   = "SNH",
  "Albury"       = "ABY",
  "Bairnsdale"   = "BDE",
  "Wyndham Vale" = "WVL",
  "Melton"       = "MEL",
  # Skybus
  "Eastern Express" = "SKY",
  "Peninsula Express" = "SKY",
  "Avalon City Express" = "AVV",
  "Melbourne City Express" = "SKY",
  "Sunshine Express" = "SKY",
  # The Overland
  "The Overland" = "OVL" 
  
)

summary[route_display %chin% names(route_bullet_changes),
        route_bullet_code := route_bullet_changes[route_display]]

# Remove brackets from route bullets
summary[
  !is.na(route_bullet_code),
  route_bullet_code := gsub(" \\([^)]*\\)", "", route_bullet_code)
]


rm(route_bullet_changes)



# Add title name with route etc
summary[, route_title_display := route_display]
summary[!is_combined_corridor & mode_type %in% c(1,2), 
        route_title_display := paste(route_title_display, "Line")]

summary[is_combined_corridor & mode_type %in% c(1,2), 
        route_title_display := paste(route_title_display, "Lines")]

summary[!is_combined_corridor & mode_type %in% c(3,4), 
        route_title_display := paste("Route", route_title_display)]

summary[is_combined_corridor & mode_type %in% c(3,4), 
        route_title_display := paste("Routes", route_title_display, "")]

summary[grep("Corridor", route_display), 
        route_title_display := route_display]

setcolorder(
  summary,
  c(
    "mode_type",
    "route_number",
    "route_display",
    "route_bullet_code",
    "route_title_display",
    "is_combined_corridor",
    "route_colour",
    "route_destination_1",
    "route_destination_0",
    "frequency_summary",
    "peak_medians",
    "span_stats"
  )
)
setkey(summary, route_number)

# Build a NAMED list keyed by route_number (object form, not array).
# Each list-column already holds one list per row, so pull [[1]] out of it
# explicitly rather than letting as.list() leave it double-wrapped.
summary_list <- lapply(
  split(summary, by = "route_number", keep.by = TRUE),
  function(dt) {
    l <- as.list(dt[1])
    l$frequency_summary <- l$frequency_summary[[1]]
    l$peak_medians      <- l$peak_medians[[1]]
    l$span_stats        <- l$span_stats[[1]]
    l$weekly_stats      <- l$weekly_stats[[1]]
    l
  }
)


if (!dir.exists("route_jsons")) dir.create("route_jsons")
if (!dir.exists("data")) dir.create("data")
# Write single combined summary JSON, keyed by route_number
jsonlite::write_json(
  summary_list,
  path       = "data/summary.json",
  auto_unbox = TRUE,
  pretty     = TRUE,
  na         = "null"
)

# Manifest
manifest <- list(routes = names(summary_list))
jsonlite::write_json(
  manifest,
  "data/manifest.json",
  auto_unbox = TRUE,
  pretty     = TRUE
)
```

## Create shape files for each route
```{r}
# Add the unshared parts of combined corridors
combine_corridors <- function(st, corridor_defs){
  # corridor_defs: named list, e.g. list("1/6" = c("1","6"), "3/64/67" = c("3","64","67"))
  # name = new combined route_short_name, value = character vector of routes to merge
  
  setkey(st, route_short_name)  # enables fast binary-search subsetting below instead of full scans
  
  # Remove existing combined corridors
  st <- st[!route_short_name %chin% names(corridor_defs)]
  
  combined_list <- lapply(names(corridor_defs), function(new_name){
  routes <- corridor_defs[[new_name]]
  subsets <- lapply(routes, function(r) st[.(r)])
  
  stop_id_lists <- lapply(subsets, function(x) x$stop_id)
  shared_stops <- Reduce(intersect, stop_id_lists)
  
  combined <- rbindlist(subsets)  # keep ALL stops from every route, not just shared
  combined[, route_short_name := new_name]
  combined
  })
  
  # Single rbind at the end, not one per corridor - avoids repeatedly copying the growing table
  rbind(st, rbindlist(combined_list))
}


stop_times <- combine_corridors(stop_times, corridor_defs)


next_stop_id <- function(stop_times){
  stop_times <- copy(stop_times)
  setorder(stop_times, route_short_name, day_number, trip_id, stop_sequence)
  
  stop_times[, next_stop_id := shift(stop_id, 1L, type = "lead"), 
             by = .(route_short_name, trip_id, day_number)]
  stop_times[,next_departure_time_m := shift(departure_time_m, 1L, type = "lead"), 
             by = .(route_short_name, trip_id, day_number)]
  stop_times[, next_shape_dist_traveled := shift(shape_dist_traveled, 1L, type = "lead"),
        by = .(route_short_name, trip_id, day_number)]
  
  stop_times[, stop_pair_id := paste0(stop_id, "-", next_stop_id, "-", direction_id)]
  stop_times[, time_m := next_departure_time_m - departure_time_m] 
  stop_times[, distance_m := next_shape_dist_traveled - shape_dist_traveled]


  unique(stop_times, by = c("route_short_name", "stop_pair_id", "day_number", "departure_time_m"))
}

stop_times_pairs  <- next_stop_id(stop_times)

stop_times_pairs <- stop_times_pairs[!is.na(next_stop_id)]

# Deal with timetables not having enough entries for each stop
# Sort descending within each group - this achieves what rev() did, but as one vectorized sort
setorder(stop_times_pairs, route_short_name, trip_id, mode_number, day_number, -stop_sequence)

# cumsum is GForce-optimized - fast even across huge numbers of groups
stop_times_pairs[, grp := cumsum(time_m != 0), by = .(route_short_name, trip_id, mode_number, day_number)]

# restore original row order
setorder(stop_times_pairs, route_short_name, trip_id, mode_number, day_number, stop_sequence)
```

```{r}
build_stop_pair_shapes <- function(stop_times, stops) {
  
  # Get unique stop pairs with their two stop IDs
  duplets <- stop_times[, .SD[1], by = "stop_pair_id"][, .(stop_pair_id, stop_id, next_stop_id, mode_number)]
  
  # Join coordinates for each stop - note averaging means to deal with multiple train stations
  stop_coords <-   stops[, .(stop_lat = mean(stop_lat), stop_lon = mean(stop_lon)), by = .(stop_id, mode_number)]
  
  duplets <- duplets[stop_coords, on = c("stop_id", "mode_number"), nomatch = 0]
  setnames(duplets, c("stop_lat", "stop_lon"), c("lat1", "lon1"))
  
  duplets <- duplets[stop_coords, on = c("next_stop_id" = "stop_id", "mode_number"), nomatch = 0]
  setnames(duplets, c("stop_lat", "stop_lon"), c("lat2", "lon2"))
  
  # Build straight-line geometries
  duplets_sf <- st_sf(
    stop_pair_id = duplets$stop_pair_id,
    displacement_m = geosphere::distHaversine(cbind(duplets$lon1, duplets$lat1), 
                                              cbind(duplets$lon2, duplets$lat2)),
    geometry = st_sfc(
      mapply(function(lon1, lat1, lon2, lat2) {
        st_linestring(rbind(c(lon1, lat1), c(lon2, lat2)))
      }, duplets$lon1, duplets$lat1, duplets$lon2, duplets$lat2,
      SIMPLIFY = FALSE),
      crs = 4326
    )
  )
  
  return(duplets_sf)
}

stop_pair_shapes <- build_stop_pair_shapes(stop_times_pairs, stops) 
# Diagnostic
stop_pairs_plot <- function(rsn){
  select_stop_pairs <- stop_times_pairs[route_short_name == rsn]$stop_pair_id
  ggplot(stop_pair_shapes[stop_pair_shapes$stop_pair_id %in% select_stop_pairs, ]) +
    geom_sf() +
    theme_minimal() +
    theme(axis.text = element_blank()) 

}

# stop_pairs_plot("742")
# stop_pairs_plot("Hurstbridge/Mernda")
```

# Speed and velocity calculation
```{r}
# Add displacement
stop_times_pairs <- as.data.table(st_drop_geometry(stop_pair_shapes))[stop_times_pairs, 
                                                                      on = "stop_pair_id"]

# Grouped metrics
stop_times_pairs[, `:=`(
  grp_time_m         = sum(time_m),
  grp_distance_m      = sum(distance_m),
  grp_displacement_m  = sum(displacement_m)
), by = .(route_short_name, trip_id, mode_number, day_number, grp)]

stop_times_pairs[, avg_speed_kmh    := fifelse(grp_time_m > 0,
                                round((grp_distance_m / 1000) / (grp_time_m / 60), 2), NA_real_)]
stop_times_pairs[, avg_velocity_kmh := fifelse(grp_time_m > 0,
                                round((grp_displacement_m / 1000) / (grp_time_m / 60), 2), NA_real_)]

```

```{r}
## Function to get the gap value that :00 or :30 falls on, whole-network, per route+stop_pair
get_30_minute_gap_snapshots <- function(stop_times_pairs){
  st <- copy(stop_times_pairs)
  
  st[, week_time := departure_time_m + day_number * 24 * 60]
  
  key_cols <- c("route_short_name", "stop_pair_id")
  
  first_buffer <- st[, .SD[which.min(week_time)], by = key_cols][, week_time := week_time - 7 * 1440]
  last_buffer  <- st[, .SD[which.max(week_time)], by = key_cols][, week_time := week_time + 7 * 1440]
  
  st <- rbind(st, first_buffer, last_buffer, fill = TRUE)
  
  setorder(st, route_short_name, stop_pair_id, week_time)
  
  st[, next_week_time := shift(week_time, type = "lead"), by = key_cols]
  st <- st[!is.na(next_week_time)]
  st[, gap := next_week_time - week_time]
  
  groups <- unique(st[, ..key_cols])
  snapshots <- groups[, CJ(snapshot_week_time = seq(0, 7 * 1440 - 30, by = 30)), by = key_cols]
  
  # Carry through the speed/velocity 
result <- st[
  snapshots,
  on = c(key_cols, "week_time <= snapshot_week_time", "next_week_time > snapshot_week_time"),
  .(route_short_name, stop_pair_id, direction_id, mode_number, gap,
    avg_speed_kmh, avg_velocity_kmh,
    trip_week_time = week_time,             # departure time of the matched trip
    snapshot_week_time = i.snapshot_week_time)
]
  
  result[, `:=`(
    day_number = snapshot_week_time %/% 1440,
    snapshot_m = snapshot_week_time %% 1440
  )]
  
  result[, gap_group := fcase(
    gap <= 10, 10, gap <= 15, 15, gap <= 20, 20,
    gap <= 30, 30, gap <= 60, 60, default = NA_real_
  )]
  
speed_buffer_min <- 30   # how close (mins) a snapshot must be to a real departure
result[, mins_since_departure := snapshot_week_time - trip_week_time]
result[mins_since_departure > speed_buffer_min,
       `:=`(avg_speed_kmh = NA_real_, avg_velocity_kmh = NA_real_)]

result[, .(route_short_name, stop_pair_id, direction_id, mode_number,
           day_number, snapshot_m, gap, gap_group, avg_speed_kmh, avg_velocity_kmh)]
}

snapshots_waits <- get_30_minute_gap_snapshots(stop_times_pairs)
snapshots_waits <- snapshots_waits[!is.na(direction_id)]


```

```{r}
# Nest gap/gap_group into direction -> day structure, per route_short_name + stop_pair_id
# Nest gap/gap_group into day -> direction structure,
# per route_short_name + stop_pair_id
nested <- snapshots_waits[
  , {
      sp <- split(.SD, by = c("direction_id", "day_number"), keep.by = FALSE)

      # Split names are "direction.day", e.g. "1.0"
      split_names <- names(sp)

      # Build direction -> day
      out <- list()

      for (i in seq_along(sp)) {

        parts <- strsplit(split_names[i], ".", fixed = TRUE)[[1]]

        direction <- parts[1]
        day <- as.integer(parts[2]) + 1L

        if (is.null(out[[direction]])) {
          out[[direction]] <- vector("list", 7)
          names(out[[direction]]) <- as.character(1:7)
        }

        x <- sp[[i]]

        out[[direction]][[day]] <- list(
          gap = x$gap,
          gap_group = x$gap_group,
          speed = x$avg_speed_kmh,
          velocity = x$avg_velocity_kmh
        )
      }

      # Give the direction lists their names
      direction_ids <- names(out)
      names(out) <- direction_ids

      # Fill missing days
      for (dir in direction_ids) {
        for (d in 1:7) {
          if (is.null(out[[dir]][[d]])) {
            out[[dir]][[d]] <- list(
              gap = rep(10080, 48),
              gap_group = rep(NA_real_, 48)
            )
          }
        }
      }

      list(data = list(out))
    },
  by = .(route_short_name, stop_pair_id)
]


# Join geometry back - built once for the whole network via build_stop_pair_shapes()
geom_lookup <- data.table(
  stop_pair_id = stop_pair_shapes$stop_pair_id,
  geometry     = st_geometry(stop_pair_shapes)
)
nested <- merge(nested, geom_lookup, by = "stop_pair_id", all.x = TRUE)
nested <- st_as_sf(nested)


for (r in unique(nested$route_short_name)) {
  route_out <- list(interactive = (nested[nested$route_short_name == r, c("stop_pair_id", "data", "geometry")]))
  
  safe_name <- gsub("/", "-", r, fixed = TRUE)  # "513/514" -> "513-514"
  
  jsonlite::write_json(
    route_out,
    path       = file.path("route_jsons", paste0(safe_name, ".json")),
    auto_unbox = TRUE,
    pretty     = FALSE,
    na         = "null"
  )
}
```

## Export Date range
```{r}
date_range <- c(first_date, last_date)
jsonlite::write_json(date_range, path = "data/date_range.json")
```

## Export .csv
```{r}
# CSV export for metrics.html "download full csv" button
mode_labels_json <- jsonlite::fromJSON("modes.json")
mode_label_lookup <- setNames(sapply(mode_labels_json, `[[`, "label"), names(mode_labels_json))

build_csv_row <- function(l) {
  freq_by_day <- if (!is.null(l$frequency_summary)) {
    setNames(l$frequency_summary, sapply(l$frequency_summary, `[[`, "day"))
  } else {
    list()
  }
  get_freq <- function(day, field) {
    v <- freq_by_day[[day]][[field]]
    if (is.null(v)) NA_real_ else v
  }

  peak   <- if (!is.null(l$peak_medians)) l$peak_medians[[1]] else list()
  weekly <- if (!is.null(l$weekly_stats)) l$weekly_stats[[1]] else list()
  get_or_na <- function(lst, field) {
    v <- lst[[field]]
    if (is.null(v)) NA_real_ else v
  }

  data.table(
    Route      = l$route_display,
    Mode       = mode_label_lookup[[as.character(l$mode_type)]],
    `Avg Frequency Mon 7-8`    = get_freq("Monday", "morning"),
    `Avg Frequency Mon 12-1`   = get_freq("Monday", "daytime"),
    `Avg Frequency Mon 12-1`        = get_freq("Monday", "evening"),
    `Morning 7-8 (Saturday)`  = get_freq("Saturday", "morning"),
    `Daytime 12-1 (Saturday)` = get_freq("Saturday", "daytime"),
    `Evening (Saturday)`      = get_freq("Saturday", "evening"),
    `Morning 7-8 (Sunday)`    = get_freq("Sunday", "morning"),
    `Daytime 12-1 (Sunday)`   = get_freq("Sunday", "daytime"),
    `Evening (Sunday)`        = get_freq("Sunday", "evening"),
    `AM Peak Median Gap (min)` = get_or_na(peak, "Median AM Peak Wait"),
    `PM Peak Median Gap (min)` = get_or_na(peak, "Median PM Peak Wait"),
    `Weekly Total Services`    = get_or_na(weekly, "Weekly Trip Count"),
    `Weekly Distance (km)`     = get_or_na(weekly, "Weekly Distance (km)"),
    `Weekly Service Hours`     = get_or_na(weekly, "Weekly Time (hr)"),
    `Average Speed (km/h)`     = get_or_na(weekly, "Average Speed (km/hr)")
  )
}

csv_dt <- rbindlist(lapply(summary_list, build_csv_row), fill = TRUE)

if (!dir.exists("Archive")) dir.create("Archive")
archive_filename <- paste0("frequencyfinder_metrics_", format(first_date, "%Y-%m-%d"), ".csv")
fwrite(csv_dt, file.path("Archive", archive_filename))       # local only, gitignored
fwrite(csv_dt, file.path("data", "metrics_latest.csv"))  # deployed
```




