---
title: "FrequencyFinder3"
output: html_document
date: "2026-05-23"
---


# Setup

## Clean Environment

```{r}
# List of items, mostly large things that are annoying to reinitialise
keep <- c("gtfs_bus", "gtfs_tram", "gtfs_train") 
# Remove everything that is not in the keep list
rm(list = setdiff(ls(), keep))
# Get rid of everything if needed
#rm(list = ls())
```

## Load Libraries

```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(sf)
library(lwgeom)
library(jsonlite)
library(gtfstools)
library(vicgtfstools) # My own repo https://github.com/adambain014/vicgtfstools 
library(tictoc) # For my own benchmarking enjoyment


shift <- data.table::shift # For the actual last time can people try and not name functions the same across dplyr and data.table
```

## Initialise the GTFS

```{r}
# Download the latest vic gtfs feed
if(!dir.exists("Temps/gtfs")){
  download_latest_vic_gtfs("Temps/gtfs")
}

# Open the tram GTFS
if(!exists("gtfs_tram")){
  gtfs_tram <- open_vic_gtfs("Temps/gtfs", "Metro_Tram")
}

# Open the bus GTFS
if(!exists("gtfs_bus")){
  gtfs_bus <- open_vic_gtfs("Temps/gtfs", "Metro_Bus")
}

# Open the train GTFS
if(!exists("gtfs_train")){
  gtfs_train <- open_vic_gtfs("Temps/gtfs", "Metro_Train")
}
```

## Create a stop times sorted by day and route
```{r}
refine_stop_times <- function(gtfs, target_day, target_route){


  route_ids <- gtfs$routes[route_short_name == target_route, route_id]
  
  possible_service_ids <- gtfs$trips[route_id %chin% route_ids, service_id]
  
  calendar <- gtfs$calendar
  
  start_date <- min(calendar$start_date)
  end_date <- max(calendar$end_date)
  
  date_list <- data.table(target_date = seq(from = start_date, to = end_date)
                          )[tolower(weekdays(target_date)) == target_day]
  
  get_date_service_ids <- function(target_date, gtfs){
    
    base_service_ids <- calendar[
      start_date <= target_date &
      end_date >= target_date &
      get(target_day) == 1,
      service_id
    ]
    
    exception_1_service_ids <- gtfs$calendar_dates[
      date == target_date & exception_type == 1,
      service_id
    ]
    
    exception_2_service_ids<- gtfs$calendar_dates[
      date == target_date & exception_type == 2,
      service_id
    ]
    
#work around for buses - I am not happy with it but due to bizarre errors in the VIC GTFS I'm going to have to do it. Didnt work
    if(gtfs$levels$mode_number[1] == 0){
      
      target_day_code <- fifelse(target_day == "sunday", "Su", fifelse(target_day == "saturday", "Sa", "MF"))
      
      calendar[,day_code := substring(service_id, 1, 2)]
      
      exception_2_service_ids <- calendar[day_code != target_day_code]$service_id
    }
    
    
    final_service_ids <- setdiff(union(base_service_ids, exception_1_service_ids), exception_2_service_ids)
    
    intersect(final_service_ids, possible_service_ids)
    
  }
  
  date_list[
    ,
    service_ids := lapply(
      target_date,
      get_date_service_ids,
      gtfs = gtfs
    )
  ]
  
  service_ids_list <- date_list[
  ,
    .(service_signature = vapply(
      service_ids,
      function(x) paste(sort(x), collapse = "|"),
      character(1)
    ),
    service_ids)
  ][    service_ids != "character(0)"
    ]
  
  best_service_signature <- service_ids_list[, .N,
    by = service_signature
  ][order(-N)][1]$service_signature

  # Just take the first one if its buses because buses are fucked in the GTFS service calendar if you look too far forward
    if(gtfs$levels$mode_number[1] == 4){
    best_service_signature <- service_ids_list[1]$service_signature
  }

  list_of_valid_service_ids <- service_ids_list[
      service_signature == best_service_signature
    ]$service_ids |> unlist() |> unique()
    
  trip_ids <- gtfs$trips[
    route_id %chin% route_ids & service_id %chin% list_of_valid_service_ids,
    trip_id
  ]
  
  # No service on this day
  if (length(trip_ids) == 0L) {
    return(data.table(
      trip_id = NA_character_,
      departure_time = NA_character_,
      stop_id = NA_character_,
      stop_sequence = NA_integer_,
      shape_dist_traveled = NA_real_,
      direction_id = NA_integer_,
      departure_time_m = NA_real_
    )[0])
  }

  #  Filter stop_times by the trip_ids
  stop_times <- gtfs$stop_times[trip_id %chin% trip_ids]

  # Keep only needed columns
  cols_to_keep <- c(
    "trip_id",
    "departure_time",
    "stop_id",
    "stop_sequence",
    "shape_dist_traveled"
  )
  stop_times <- stop_times[, ..cols_to_keep]
  
  # Add direction_id
  direction_ids <- gtfs$trips[trip_id %chin% trip_ids, .(trip_id, direction_id)]
  stop_times <- stop_times[direction_ids, on = "trip_id"]
  
  # Turn departure times into seconds
  departure_times_to_minutes <- function(dt) {
    dt <- copy(dt)
    dt[, c("h","m","s") := tstrsplit(departure_time, ":", type.convert = TRUE)]
    dt[, departure_time_m := round(h*60 + m + s/60)]
    dt[, !c("h","m","s")]
  }

  stop_times <- departure_times_to_minutes(stop_times)
  
  # remove dupes
  stop_times <- stop_times[, .SD[1], by = .(stop_id, departure_time_m, direction_id, shape_dist_traveled)]
  
  return(stop_times)
}


test <- gtfs_bus$trips[,.(trip_id, service_id)][gtfs_bus$stop_times[trip_id %chin% gtfs_bus$trips[route_id %chin% gtfs_bus$routes[route_short_name == "742"]$route_id]$trip_id][,.SD[1], by = .(departure_time, stop_id)][order(departure_time)][stop_id == 15643], on = "trip_id"] |> setcolorder("service_id", 1)

#19-742--1-MF11-32

test <- gtfs_bus$calendar[service_id %chin% gtfs_bus$trips[route_id %chin% gtfs_bus$routes[route_short_name == "742"]$route_id]$service_id]

test <- refine_stop_times(gtfs_bus, "monday", 742)[stop_id == 15643][order(departure_time_m)]

print(test)
```

# Summary Stats
Average Frequency for Morning, Daytime, and Evening across Monday, Saturday and Sunday
```{r}
# Function to add minutes until next service in same direction at same stop
time_until_next_service <- function(stop_times){
  
  waits <- copy(stop_times)
  
  # Sort to ensure correct order
  setorder(waits, stop_id, direction_id, departure_time_m)
  
  waits[,
    next_time_m := shift(departure_time_m, 1L, type = "lead"),
    by = .(stop_id, direction_id)
  ]
  
  # Drop last departure per group (no next service) — NAs now land correctly
  waits <- waits[!is.na(next_time_m)]
  
  waits[, wait_m := next_time_m - departure_time_m]
  
  return(waits)
}

freq_at_window <- function(waits, start_min, end_min) {
  window_waits <- waits[departure_time_m >= start_min & departure_time_m < end_min]
  
  if (nrow(window_waits) == 0) return(NA_real_)
  
  # Mean wait per stop+direction, then average across all stop+direction combos
  mean_wait <- window_waits[, .(mean_wait = mean(wait_m)), 
                              by = .(stop_id, direction_id)][, mean(mean_wait)]
  
  # Round to nearest 0.5
  mean_wait <- round(mean_wait*2)/2
  
  return(mean_wait)
}

# Function to derive wait times
freq_at_3_windows <- function(waits){

  morning <- freq_at_window(waits, 7 * 60, 8 * 60)
  daytime <- freq_at_window(waits, 12 * 60, 13 * 60)
  evening <- freq_at_window(waits, 21 * 60, 22 * 60)

  out <- data.table(morning = morning,
                    daytime = daytime,
                    evening = evening)
  }

# Test usage
stop_times_monday   <- refine_stop_times(gtfs_bus, "monday", "742")
stop_times_saturday <- refine_stop_times(gtfs_bus, "saturday", "742")
stop_times_sunday   <- refine_stop_times(gtfs_bus, "sunday", "742")

waits_monday   <- time_until_next_service(stop_times_monday)
waits_saturday <- time_until_next_service(stop_times_saturday)
waits_sunday   <- time_until_next_service(stop_times_sunday)

freq_monday   <- freq_at_3_windows(waits_monday)[, day := "Monday"]
freq_saturday <- freq_at_3_windows(waits_saturday)[, day := "Saturday"]
freq_sunday   <- freq_at_3_windows(waits_sunday)[, day := "Sunday"]

frequency_summary <- rbind(freq_monday, freq_saturday, freq_sunday) |>
  setcolorder("day", 1)
```

Peak medians
```{r}
peak_median <- function(waits, AMPM) {
  
  if (AMPM == "AM") {
    start_min <- 7 * 60   # 6:00
    end_min   <- 9 * 60   # 9:00
  }
  
  if (AMPM == "PM") {
    start_min <- 16 * 60  # 4:00 PM
    end_min   <- 19 * 60  # 7:00 PM
  }
  
  # Filter to the peak window
  window <- waits[
    departure_time_m >= start_min &
    departure_time_m <  end_min
  ]
  
  if (nrow(window) == 0) return(NA_real_)
  
  median(window$wait_m)
}

take_peak_medians <- function(waits){
  
  peak_medians <- data.table(`Median AM Peak Wait` = peak_median(waits, "AM"),
                             `Median PM Peak Wait` = peak_median(waits, "PM"))
  
  return(peak_medians)
}

# Test usage
stop_times_monday   <- refine_stop_times(gtfs_bus, "monday", "907")
waits_monday        <- time_until_next_service(stop_times_monday)
peak_medians        <- take_peak_medians(waits_monday)
```

Span stats
From Monday - Friday - Saturday - Sunday to pick up night network

```{r}
# Create a function to make a big unified stop times block across week
unified_stop_times <- function(gtfs, target_route){
  
  stop_times_monday    <- refine_stop_times(gtfs, "monday",    target_route)[,day := 1]
  stop_times_tuesday   <- refine_stop_times(gtfs, "tuesday",   target_route)[,day := 2]
  stop_times_wednesday <- refine_stop_times(gtfs, "wednesday", target_route)[,day := 3]
  stop_times_thursday  <- refine_stop_times(gtfs, "thursday",  target_route)[,day := 4]
  stop_times_friday    <- refine_stop_times(gtfs, "friday",    target_route)[,day := 5]
  stop_times_saturday  <- refine_stop_times(gtfs, "saturday",  target_route)[,day := 6]
  stop_times_sunday    <- refine_stop_times(gtfs, "sunday",    target_route)[,day := 7]
  
  stop_times_all <- rbind(stop_times_monday, 
                          stop_times_tuesday, 
                          stop_times_wednesday, 
                          stop_times_thursday, 
                          stop_times_friday, 
                          stop_times_saturday, 
                          stop_times_sunday)
  
  # Mark for past midnight
  stop_times_all <- stop_times_all[,past_midnight := departure_time_m >= 24 * 60]
  # Move to the next day
  #stop_times_all <- stop_times_all[past_midnight == TRUE, day := day + 1]
  #stop_times_all <- stop_times_all[day == 8, day := 1]
  # Change 25 to 1 etc.
  #stop_times_all <- stop_times_all[past_midnight == TRUE, departure_time_m := departure_time_m - 24 * 60]
  
  return(stop_times_all)
}


    # Guard against infinites caused by missing days
  safe_max <- function(x) if (length(x) == 0 || all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
  safe_min <- function(x) if (length(x) == 0 || all(is.na(x))) NA_real_ else min(x, na.rm = TRUE)

# Function to break one day into spans
one_day_span <- function(unified_stop_times_o, target_day_number){
  
  day_stop_times <- unified_stop_times_o[day == target_day_number]
 
  # Remove terminus stops from consideration due to multiple bays/platofrms being randomly used 
 # day_stop_times <- day_stop_times[, .SD[-c(1, .N)], by = trip_id]
  
  valid_stops <- day_stop_times[,.N, by = "stop_id"][,pc_max := N / safe_max(N)][pc_max > 0.2]$stop_id
  
  day_stop_times <- day_stop_times[stop_id %chin% valid_stops]
  
  day_stop_times_arranged <- day_stop_times[order(departure_time_m),.(departure_time_m, stop_id)]
  day_stop_times_arranged <- unique(day_stop_times_arranged)
  
  span_start <- safe_min(day_stop_times_arranged$departure_time_m)
  span_end   <- safe_max(day_stop_times_arranged$departure_time_m)
  
  full_coverage_start <- day_stop_times_arranged[, .(first_service = safe_min(departure_time_m)), by = stop_id
                                                 ][, safe_max(first_service)]
  full_coverage_end   <- day_stop_times_arranged[, .(last_service  = safe_max(departure_time_m)), by = stop_id
                                                 ][, safe_min(last_service)]
  
    out <- data.table(
    day = target_day_number,
    span_start = span_start,
    span_end = span_end,
    full_coverage_start = full_coverage_start,
    full_coverage_end = full_coverage_end)
    
    return(out)
}

span_across_week <- function(unified_stop_times_o){
  
  monday_span    <- one_day_span(unified_stop_times_o, 1)[,day := "monday"]
  tuesday_span   <- one_day_span(unified_stop_times_o, 2)[,day := "tuesday"]
  wednesday_span <- one_day_span(unified_stop_times_o, 3)[,day := "wednesday"]
  thursday_span  <- one_day_span(unified_stop_times_o, 4)[,day := "thursday"]
  friday_span    <- one_day_span(unified_stop_times_o, 5)[,day := "friday"]
  saturday_span  <- one_day_span(unified_stop_times_o, 6)[,day := "saturday"]
  sunday_span    <- one_day_span(unified_stop_times_o, 7)[,day := "sunday"]
  
  out <- rbind(monday_span,
               tuesday_span,
               wednesday_span,
               thursday_span,
               friday_span,
               saturday_span,
               sunday_span)
  
  setcolorder(out, "day", 1)
  
  # Deal with night network
  
  out[,span_end_n                 := span_end - 24 * 60]
  out[, prior_span_end_n          := span_end_n[c(.N, 1:(.N-1))]]
  
  out[, span_diff            := prior_span_end_n - span_start]
  out[, begins_night_network := span_diff >= 30]
  out[, ends_night_network   := begins_night_network[c(2:.N, 1)]]
  
  out[begins_night_network == TRUE, span_start := 3 * 60]
  out[begins_night_network == TRUE, full_coverage_start := 3 * 60]
  
  out[ends_night_network == TRUE, span_end := 27 * 60]
  out[ends_night_network == TRUE, full_coverage_end := 27 * 60]
    
  # Simplify it a bit
  out <- out[,.(day, span_start, span_end, full_coverage_start, full_coverage_end)]
  
  
  return(out)
  
}
  
# Test

week_stop_times <- unified_stop_times(gtfs_bus, "150")
test <- one_day_span(week_stop_times, 7)
span_summary <- span_across_week(week_stop_times)
```

# Example Plot
```{r}
plot_week_spans <- function(spans) {
  
  # Build long format for all days
  segments <- spans[, .(
    type  = c("Span",  "Full Coverage"),
    start = c(span_start, full_coverage_start),
    end   = c(span_end, full_coverage_end)
  ), by = day]
  
  
  # Convert to hours
  segments[, start_h := start / 60]
  segments[, end_h   := end   / 60]
  
  # Collapse midnight into same type as its counterpart for colour
  segments[, type_colour := fcase(
    type == "Span",          "Span",
    type == "Full Coverage", "Full Coverage"
  )]
  
  # Order days
  day_levels <- c("monday","tuesday","wednesday","thursday","friday","saturday","sunday")
  day_labels <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
  segments[, day := factor(day, levels = day_levels, labels = day_labels)]
  
  ggplot(segments, aes(y = type_colour, colour = type_colour, group = type)) +
    geom_segment(aes(x = start_h, xend = end_h, yend = type_colour), linewidth = 4) +
    scale_x_continuous(
      breaks = seq(3, 27, by = 3),
      labels = function(x) sprintf("%02d:00", as.integer(x)),
      limits = c(3, 27)
    ) +
    scale_colour_manual(
      values = c("Span" = "#2196F3", "Full Coverage" = "#FF5722"),
      name   = NULL
    ) +
    facet_wrap(~ day, ncol = 1, strip.position = "left") +
    labs(
      title = "Service span across the week",
      x = NULL,
      y = NULL
    ) +
    theme_minimal() +
    theme(
      strip.text.y.left = element_text(angle = 0, hjust = 1),
      axis.text.y       = element_blank(),
      axis.ticks.y      = element_blank(),
      legend.position   = "bottom"
    )
}

# Usage
week_stop_times <- unified_stop_times(gtfs_bus, "150")
span_summary <- span_across_week(week_stop_times)
plot_week_spans(span_summary)
```

## Leaflet interactive
Split days into 30 min chunks across 7 days of the week, and assign a frequency to each "duplet"

First lets make duplet sfs

```{r}
next_stop_id <- function(stop_times){
  stop_times <- as.data.table(stop_times)
  setorder(stop_times, day, trip_id, stop_sequence)
  stop_times <- stop_times[, next_stop_id := shift(stop_id, 1L, type = "lead"), by = .(trip_id, day)
                              ][, next_shape_dist_traveled := shift(shape_dist_traveled, 1L, type = "lead"), 
                                by = .(trip_id, day)
                              ][, stop_pair_id := paste0(stop_id, "-", next_stop_id, "-", direction_id)
                              ][!is.na(next_stop_id)
                              ]
  
  unique(stop_times, by = c("stop_pair_id", "day", "departure_time_m"))
}

stop_times_week <- unified_stop_times(gtfs_tram, 1)
stop_times_pairs  <- next_stop_id(stop_times_week)

build_stop_pair_shapes <- function(stop_times, gtfs) {
  
  # Get unique stop pairs with their two stop IDs
  duplets <- stop_times[, .SD[1], by = "stop_pair_id"][, .(stop_pair_id, stop_id, next_stop_id)]
  
  # Join coordinates for each stop
  stop_coords <- gtfs$stops[, .(stop_id, stop_lat, stop_lon)]
  
  duplets <- duplets[stop_coords, on = "stop_id", nomatch = 0]
  setnames(duplets, c("stop_lat", "stop_lon"), c("lat1", "lon1"))
  
  duplets <- duplets[stop_coords, on = c("next_stop_id" = "stop_id"), nomatch = 0]
  setnames(duplets, c("stop_lat", "stop_lon"), c("lat2", "lon2"))
  
  # Build straight-line geometries
  duplets_sf <- st_sf(
    stop_pair_id = duplets$stop_pair_id,
    geometry = st_sfc(
      mapply(function(lon1, lat1, lon2, lat2) {
        st_linestring(rbind(c(lon1, lat1), c(lon2, lat2)))
      }, duplets$lon1, duplets$lat1, duplets$lon2, duplets$lat2,
      SIMPLIFY = FALSE),
      crs = 4326
    )
  )
  
  return(duplets_sf)
}



stop_times_week <- unified_stop_times(gtfs_bus, 273)
stop_times_pairs  <- next_stop_id(stop_times_week)
stop_pair_shapes <- build_stop_pair_shapes(stop_times_pairs, gtfs_bus)

plot(stop_pair_shapes)
```

Find out current wait period for each first stop of each duplet - at every 30 minute intervals 7 days a week



```{r}
# Function to get the gap value that :00 or :30 falls on per stop_pair
get_30_minute_gap_snapshots_per_stop_id <- function(stop_times){
  stop_times_i <- stop_times[] # trim
  
  # explode into week minutes
  stop_times_i[, week_time := departure_time_m + (day - 1L) * 24 * 60] 
  
  # Fix the circular week problem by adding buffers
  first_buffer <- stop_times_i[
      , .SD[1], by = stop_pair_id
    ][
      , week_time := week_time - 7 * 1440
    ]

  last_buffer <- stop_times_i[
      , .SD[.N], by = stop_pair_id
    ][
      , week_time := week_time + 7 * 1440
    ]
  
  stop_times_i <- rbind(stop_times_i, first_buffer, last_buffer)
  
  # Order chronologically within each stop pair
  setorder(stop_times_i, stop_pair_id, week_time)
  
  # Next departure from same stop pair
  stop_times_i[
      ,
      next_week_time := shift(week_time, type = "lead"),
      by = stop_pair_id
    ]
  
  # Remove final observation (no next departure)
  stop_times_i <- stop_times_i[
      !is.na(next_week_time)
    ]
  
  # Gap between departures
  stop_times_i[
    ,
    gap := next_week_time - week_time
  ]
  
  # Generate snapshots every 30 minutes for entire week
  snapshots <- CJ(
    stop_pair_id = unique(stop_times_i$stop_pair_id),
    snapshot_week_time = seq(
      0,
      7 * 1440 - 30,
      by = 30
    )
  )
  
  # Find which interval each snapshot falls inside
  result <- stop_times_i[
    snapshots,
    on = .(
      stop_pair_id,
      week_time           <= snapshot_week_time,
      next_week_time      >  snapshot_week_time
    ),
    .(stop_pair_id, direction_id, gap, snapshot_week_time = i.snapshot_week_time)
  ]
  
    # Convert back to day/time
    result[
      ,
      `:=`(
        day        = (snapshot_week_time %/% 1440) + 1L,
        snapshot_m = snapshot_week_time %% 1440
      )
    ]
  
    # Frequency bands
    result[
      ,
      gap_group := fcase(
        gap <= 10, 10,
        gap <= 15, 15,
        gap <= 20, 20,
        gap <= 30, 30,
        gap <= 60, 60,
        default = NA_real_
      )
    ]
  
    result <- result[
      ,
      .(
        stop_pair_id,
        direction_id,
        day,
        snapshot_m,
        gap,
        gap_group
      )
    ]
            }


# test

stop_times_week <- unified_stop_times(gtfs_tram, 35)
stop_times_pairs  <- next_stop_id(stop_times_week)
stop_pair_shapes <- build_stop_pair_shapes(stop_times_pairs, gtfs_tram)

 
snapshots_waits <- get_30_minute_gap_snapshots_per_stop_id(stop_times_pairs)

snapshots_waits <- snapshots_waits[!is.na(direction_id)]


final_int <- snapshots_waits[
  ,
  .(data = list(
    lapply(split(.SD, by = "direction_id", keep.by = FALSE), function(d) {
      lapply(split(d, by = "day", keep.by = FALSE), function(dd) {
        list(gap = dd$gap, gap_group = dd$gap_group)
      })
    })
  )),
  by = stop_pair_id
]

# Join geometry back as sf object
geom_lookup <- data.frame(
  stop_pair_id = stop_pair_shapes$stop_pair_id,
  geometry     = st_geometry(stop_pair_shapes)
)
final_int <- merge(final_int, geom_lookup, by = "stop_pair_id", all.x = TRUE)
final_int <- st_as_sf(final_int)
    
```

GTFS (R pipeline)
->
precomputed_data.json
->
GitHub Pages frontend
->
JS renders maps + charts



## Function to make json dataset for each route
```{r}
make_pre_json <- function(gtfs, route_number){
  
  mode_type <- gtfs$levels$mode_number[1]
  
  # Get Colour
  colour <- gtfs$routes[route_short_name == route_number]$route_color[1]
  
  # Get destinations for directions
  route_destination_1 <- gtfs$trips[route_id %chin% gtfs$routes[route_short_name == route_number]$route_id
                                    ][direction_id == 1][,.N, by = "trip_headsign"][order(-N)
                                                                                    ][1]$trip_headsign
  route_destination_0 <- gtfs$trips[route_id %chin% gtfs$routes[route_short_name == route_number]$route_id
                                    ][direction_id == 0][,.N, by = "trip_headsign"][order(-N)
                                                                                    ][1]$trip_headsign

  # Frequency Summary
  stop_times_monday   <- refine_stop_times(gtfs, "monday", route_number)
  stop_times_saturday <- refine_stop_times(gtfs, "saturday", route_number)
  stop_times_sunday   <- refine_stop_times(gtfs, "sunday", route_number)

  waits_monday   <- time_until_next_service(stop_times_monday)
  waits_saturday <- time_until_next_service(stop_times_saturday)
  waits_sunday   <- time_until_next_service(stop_times_sunday)

  freq_monday   <- freq_at_3_windows(waits_monday)[, day := "Monday"]
  freq_saturday <- freq_at_3_windows(waits_saturday)[, day := "Saturday"]
  freq_sunday   <- freq_at_3_windows(waits_sunday)[, day := "Sunday"]

  frequency_summary <- rbind(freq_monday, freq_saturday, freq_sunday)

  # Peak Medians
  peak_medians <- take_peak_medians(waits_monday)
  
  # Span
  
  week_stop_times <- unified_stop_times(gtfs, route_number)
  span_summary <- span_across_week(week_stop_times)
  
  # Interactive
  stop_times_pairs  <- next_stop_id(week_stop_times)
    
    
  
  snapshots_waits <- get_30_minute_gap_snapshots_per_stop_id(stop_times_pairs)
  snapshots_waits <- snapshots_waits[!is.na(direction_id)]
  
  
  final_int <- snapshots_waits[
    ,
    .(data = list(
      lapply(split(.SD, by = "direction_id", keep.by = FALSE), function(d) {
        lapply(split(d, by = "day", keep.by = FALSE), function(dd) {
          list(gap = dd$gap, gap_group = dd$gap_group)
        })
      })
    )),
    by = stop_pair_id
  ]
  
  # Join geometry back as sf object
  
  stop_pair_shapes <- build_stop_pair_shapes(stop_times_pairs, gtfs)
  geom_lookup <- data.frame(
    stop_pair_id = stop_pair_shapes$stop_pair_id,
    geometry     = st_geometry(stop_pair_shapes)
  )
  final_int <- merge(final_int, geom_lookup, by = "stop_pair_id", all.x = TRUE)
  final_int <- st_as_sf(final_int)
      
        


  # FINAL STRUCTURE 
  interactive_out <- list(
    interactive = as.data.frame(final_int)
  )
  
  summary_out <- list(
    mode_type           = as.character(mode_type),
    route_number        = as.character(route_number),
    route_colour        = as.character(colour),
    route_destination_1 = as.character(route_destination_1),
    route_destination_0 = as.character(route_destination_0),
    frequency_summary   = as.data.frame(frequency_summary),
    peak_medians        = as.data.frame(peak_medians),
    span_stats          = as.data.frame(span_summary)
  )
  
  return(list(summary = summary_out, interactive = interactive_out))
}

export_json <- function(gtfs, route_number){
    
  print(paste0("Exporting ", route_number, ".json"))
  
  out <- make_pre_json(gtfs, route_number)
  
  path <- "route_jsons"
  if (!dir.exists(path)) dir.create(path, recursive = TRUE)
  
  # Write per-route interactive JSON
  jsonlite::write_json(
    out$interactive,
    path      = file.path(path, paste0(route_number, ".json")),
    auto_unbox = TRUE,
    pretty    = TRUE,
    na        = "null"
  )
  
  invisible(out$summary)
}


```

```{r}
valid_train_routes <- unique(gtfs_train$routes$route_short_name)

valid_tram_routes <- unique(gtfs_tram$routes$route_short_name)

valid_bus_routes <- unique(gtfs_bus$routes[nchar(route_short_name) == 3]$route_short_name)
#valid_bus_routes <- c("742", "903", "273", "907", "150")

# Export interactives and collect summaries
all_summaries <- c(
    #lapply(valid_train_routes,  function(r) export_json(gtfs_train,  r)),
  lapply(valid_tram_routes, function(r) export_json(gtfs_tram, r))
  ,lapply(valid_bus_routes,  function(r) export_json(gtfs_bus,  r))
)
names(all_summaries) <- c(#valid_train_routes,
                          valid_tram_routes, 
                          valid_bus_routes
                          )

# Write single combined summary JSON
jsonlite::write_json(
  all_summaries,
  path       = "route_jsons/summary.json",
  auto_unbox = TRUE,
  pretty     = TRUE,
  na         = "null"
)

# Manifest
routes <- tools::file_path_sans_ext(
  list.files("route_jsons", pattern = "\\.json$")
)
jsonlite::write_json(
  list(routes = setdiff(routes, c("manifest", "summary"))),
  "route_jsons/manifest.json",
  auto_unbox = TRUE
)

valid_bus_routes
```